---
title: "p8105_hw3_rs4025"
author: "Rebecca Silva"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

library(tidyverse)
library(viridis)
library(p8105.datasets)
library(kableExtra)
```

## Problem 1 

write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

The `instacart` dataset gives infomation about online grocery shopping in 2017 using the grocery service, Instacart. It has `r nrow(instacart)` observations of `r ncol(instacart)` variables, 4 of which are character variables, and 11 of which are integer variables. Some key variables (that we will work on) are as follows: 
* `order_id`: order number 
* `product_id`: product number
* `user_id`: user number
* `order_number`: sequence number of order for user
* `product_name` (character): name/type of product
* `aisle_id`: aisle_id
* `aisle` (character): sorts of products aisle contains 
* `department` (character): name of department

(could list first the ones that are self explanatory?)

## note order of key vars

## give illstrative examples of observations ?

To give an example of the dataset, the first 10 observations are shown below in Table 1. 
```{r}
head = head(instacart, 10) 

knitr::kable(head, 
             caption = "Table 1: First 10 Observation of Instacart dataset")
```


How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisle = 
  instacart %>% 
  group_by(aisle, aisle_id) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

There are `r nrow(aisle)` aisles in the online grocery store, the most popular of which are fresh vegetables and fresh fruits.

# remove table 
```{r}
knitr::kable(head(aisle,3), 
             caption = "Table 2: Top 3 Most Popular Aisles")
```
confused about what each separate observation is 

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
## do not know what he wants (cld order by department? no)
```{r}
aisle %>% 
  filter( n > 10000) %>% 
  ggplot(aes(x = aisle_id, fill = aisle)) + 
  geom_histogram() + 
  labs(
    title = "Number of Items by Aisle",
    x = "Aisle Number",
    y = "Number of Items")


aisle %>% 
  filter( n > 10000) %>% 
  mutate(n_rank = min_rank(desc(n))) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(
    title = "Number of Items by Aisle",
    x = "Aisle Number",
    y = "Number of Items")

```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
table_aisle = 
  instacart %>% 
  filter(aisle %in% c("baking ingredients",
                      "dog food care", 
                      "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(n = n())

# use min_rank()
  
  
  
```

