---
title: "p8105_hw3_rs4025"
author: "Rebecca Silva"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

library(tidyverse)
library(viridis)
library(p8105.datasets)
library(kableExtra)
```

## Problem 1 

write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

The `instacart` dataset gives infomation about online grocery shopping in 2017 using the grocery service, Instacart. It has `r nrow(instacart)` observations of `r ncol(instacart)` variables, 4 of which are character variables, and 11 of which are integer variables. Some key variables (that we will work on) are as follows: 
* `order_id`: order number 
* `product_id`: product number
* `user_id`: user number
* `order_number`: sequence number of order for user
* `product_name` (character): name/type of product
* `aisle_id`: aisle_id
* `aisle` (character): sorts of products aisle contains 
* `department` (character): name of department

(could list first the ones that are self explanatory?)

## note order of key vars

## give illstrative examples of observations ?

To give an example of the dataset, the first 10 observations are shown below in Table 1. 
```{r}
head = head(instacart, 10) 

knitr::kable(head, 
             caption = "Table 1: First 10 Observation of Instacart dataset")
```


How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisle = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

There are `r nrow(aisle)` aisles in the online grocery store, the most popular of which  are shown below with the number of inndividual orders(?) 

confused about what each separate observation is 