---
title: "p8105_hw3_rs4025"
author: "Rebecca Silva"
date: "10/4/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

#scale_colour_discrete = scale_colour_viridis_d
#scale_fill_discrete = scale_fill_viridis_d
#theme_set(theme_minimal() + theme(legend.position = "bottom"))

library(tidyverse)
library(viridis)
library(p8105.datasets)
library(kableExtra)
```

# Questions: 
+ delete/renumber tables 
+ plot for prob 1?


## Problem 1 

write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

The `instacart` dataset gives infomation about online grocery shopping in 2017 using the grocery service, Instacart. It has `r nrow(instacart)` observations of `r ncol(instacart)` variables, 4 of which are character variables, and 11 of which are integer variables. Some key variables (that we will work on) are as follows: 

* `order_id`: order number 
* `product_id`: product number
* `user_id`: user number
* `order_number`: sequence number of order for user
* `product_name` (character): name/type of product
* `aisle_id`: aisle_id
* `aisle` (character): sorts of products aisle contains 
* `department` (character): name of department

(could list first the ones that are self explanatory?)

## note order of key vars

## give illstrative examples of observations ?

To give an example of the dataset, the first 3 observations are shown below in Table 1. 
```{r}
head = head(instacart, 3) 

knitr::kable(head, 
             caption = "Table 1: First 10 Observation of Instacart dataset")
```


How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisle = 
  instacart %>% 
  group_by(aisle) %>%   ## dont think i need aisle_id
  summarize(n = n()) %>% 
  arrange(desc(n))
```

There are `r nrow(aisle)` aisles in the online grocery store, the most popular of which are fresh vegetables, fresh fruits, and packaged vegetables and fruit.

```{r}
knitr::kable(head(aisle,3), 
             caption = "Table 2: Top 3 Most Popular Aisles")
```
confused about what each separate observation is 

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
## do not know what he wants (cld order by department? no)
```{r}
aisle %>% 
  filter( n > 10000) %>% 
  ggplot(aes(x = aisle_id, fill = aisle)) + 
  geom_histogram() + 
  labs(
    title = "Number of Items by Aisle",
    x = "Aisle Number",
    y = "Number of Items")


## the correct one?
aisle %>% 
  ungroup() %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n)) %>% 
  filter( n > 10000) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(
    title = "Number of Items by Aisle",
    x = "Aisle",
    y = "Number of Items") +
  theme(axis.text.x = element_text(angle = 90))

## try faceting by department -- not pretty
instacart %>% 
  group_by(department, aisle) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n)) %>% 
  filter( n > 10000) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  facet_grid(. ~ department) + 
  labs(
    title = "Number of Items by Aisle",
    x = "Aisle",
    y = "Number of Items") +
  theme(axis.text.x = element_text(angle = 90))

```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
table_aisle = 
  instacart %>% 
  filter(aisle %in% c("baking ingredients",
                      "dog food care", 
                      "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(n = n()) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>% 
  select(-rank)
        
knitr::kable(table_aisle, 
            caption = "Most popular items from Baking Ingredients, Dog Food Care, and Packaged Vegetables and Fruit")

```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

note: I rounded mean hour to closest hour
#Q: still dont know if I should leave order_dow as is or covert to day of week w mutate (code is b) 
```{r}
mean_table = 
  instacart %>% 
  filter( product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select( product_name, order_dow, order_hour_of_day) %>% 
  group_by( product_name, order_dow) %>% 
  summarize(mean = round(mean(order_hour_of_day))) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean)

#mutate(order_dow = recode(as.character(order_dow), `1` = "Monday", `2` = "Tuesday")) %>%

knitr::kable(mean_table, 
            caption = "Mean Hour of Day for Orders of Pink Lady Apples and Coffee Ice Cream")
```





## Problem 2 